Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions

### 2024-07-24 latex/intro.tex
Moved laplacian like discussion to intro

The inverse problem of the 1D heat equation is difficult ("ill-posed")
since heat spreads in a diffusive manner and any roughness in the
initial condition is quickly smoothed, see Supplementary movies S1 and
S2 (code generating these movies is located in module
\texttt{movies.py} in the accompanying
\href{https://github.com/yairdaon/OED}{repository}). For ill-posed
problems regularization is required, and we implement such
regularization via a fully Bayesian formulation of the inverse
problem, see Section \ref{section:prelims} for details.

>>>>

The inverse problem of the 1D heat equation is difficult ("ill-posed")
since heat spreads in a diffusive manner and any roughness in the
initial condition is quickly smoothed, see Supplementary movies S1 and
S2 (code generating these movies is located in module
\texttt{movies.py} in the accompanying
\href{https://github.com/yairdaon/OED}{repository}). For ill-posed
problems regularization is required, and we implement such
regularization via a fully Bayesian formulation of the inverse
problem.

Specifically, we set a prior on the initial condition $u_0 \sim
\mathcal{N}(0, (-\Delta)^{-1})$, where we impose the homogeneous
Dirichlet boundary condition on the Laplacian. We choose
$(-\Delta)^{-1}$ as our covariance operator following \cite{Stuart10}:
our choice ensures the power of the negative Laplacian inside the
parentheses is larger than $d/2$, where $d=1$ is the dimension of our
computational domain $\Omega$. One can intuitively understand this
choice from Weyl's asymptotics, that states that the eigenvalues of
Laplacian increase as $n^{d/2}$. Thus, our choice of covariance
$(-\Delta)^{-1}$ ensures higher modes decay sufficiently quickly, so
prior samples have sufficient regularity. Our choice of prior
covariance ensures the posterior is well-defined and well-behaved
\cite[Theorem 3.1 and Lemma 6.25]{Stuart10}.

### 2024-07-24 latex/prelims.tex
Moved laplacian like discussion to intro

We take a Gaussian prior \(\param \sim \pr = \normal(\prmean
,\prcov)\) with some appropriate covariance operator \(\prcov\) on
\(\hilp\) \cite{Stuart10}. For the inverse problem of the 1D heat
equation we choose $\prcov = -\Delta^{-1}$ --- see Section
\ref{subsec:1d heat} for details and a short discussion of
"Laplacian-like" covariance operators. Note that \(\fwd \prcov
\fwd^*\) is the prior covariance in \(\hilo\) \cite{Stuart10}, and as
such is assumed invertible --- an assumption which we will use later
(if \(\fwd\) has a nontrivial kernel we utilize Occam's Razor and
ignore said kernel altogether).

Since $\fwd$ is linear and $\pr$ is Gaussian --- the posterior
\(\post\) is Gaussian as well. We do not utilize the posterior mean in
this study, but the posterior covariance operator $\postcov$ is given
by \cite{Stuart10}:
\begin{align}\label{eq:postcov}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
  )^{-1}.
\end{align}

>>>>

We take a Gaussian prior \(\param \sim \pr = \normal(\prmean
,\prcov)\) with some appropriate covariance operator \(\prcov\) on
\(\hilp\) \cite{Stuart10}. For example, for the inverse problem of the
1D heat equation we chose $\prcov = -\Delta^{-1}$ --- see Section
\ref{subsec:toy}. Note that \(\fwd \prcov \fwd^*\) is the prior
covariance in \(\hilo\) \cite{Stuart10}, and as such is assumed
invertible --- an assumption which we will use later (if \(\fwd\) has
a nontrivial kernel we utilize Occam's Razor and ignore said kernel
altogether).

Since $\fwd$ is linear and $\pr$ is Gaussian --- the posterior
\(\post\) is Gaussian as well. We do not utilize the posterior mean in
this study, but the posterior covariance operator $\postcov$ is given
by the known formula\cite{Stuart10}:
\begin{align}\label{eq:postcov}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
  )^{-1}.
\end{align}
